---
title: "Amazon Spending"
output:
  html_notebook: default
  pdf_document: default
  word_document: default
---

```{r set working directory}
setwd("C:/Users/Branch/Documents/misc. code and homework/DSC680/Amazon project")
getwd()
```

```{r tidyverse includes ggplot2, dplyr, tidyr, and readr, purrr, tibble, stringr, forcats}
install.packages("tidyverse")
library(tidyverse)
install.packages("lubridate")
library(lubridate)
install.packages("plotly")
library(plotly)
install.packages("moderndive")
library(moderndive)
install.packages("skimr")
library(skimr)
install.packages("ISLR")
library(ISLR)
install.packages("ggplot2")
library(ggplot2)
# The ones below here are for ARIMA
install.packages("rio")
install.packages("forecast")
install.packages("tseries")
library(rio)
library(forecast)
library(tseries)
install.packages("anytime")
library(anytime)
```


```{r import first dataset after combining self and spouse order history in Excel and cleaning}
data <- read.csv('cl_Amazon_both.csv',stringsAsFactors = FALSE)
head(data)
```

```{r delete unecessary columns}
df = subset(data, select = -c(Website, OriginalListPrice))
head(df)
```

```{r exploration}
summary(df)
```

```{r can use str() if glimpse() does not work}
str(df)
```

```{r check class of currency variable(s)}
class(df$ListPriceCorrected)
```
# from https://stackoverflow.com/questions/16384933/create-an-id-row-number-column
```{r add index column}
df$ID <- 1:nrow(df)
head(df)
```

# from https://stackoverflow.com/questions/22286419/move-a-column-to-first-position-in-a-data-frame
```{r move ID from last column to first}
df <- df[,c(31, 1:30)]
head(df)
```

```{r see entire dataframe on another tab}
View(df)
```

# from https://www.earthdatascience.org/courses/earth-analytics/time-series-data/summarize-time-series-by-month-in-r/
# If this plot doesn't work, check the ggplot2 box under packages.  Has to do with newer version of R.
```{r plot the data using ggplot2 and pipes}
ggplot(df, aes(x = OrderDate, y = ItemTotal)) +
    geom_point(color = "darkorchid4") +
    labs(title = "Amazon Spending by One Family",
         subtitle = "Includes purchases made with Amazon rewards",
         y = "Amount Spent",
         x = "Order Date (June 2006 - July 2021)") + theme_bw(base_size = 15)
```

```{r add facet charts, like grouping by a variable}
ggplot(df, aes(x = OrderDate, y = ItemTotal)) +
    geom_point(color = "darkorchid4") +
    facet_wrap(~ Year) +
    labs(title = "Family Spending on Amazon Across the Year",
         subtitle = "Includes purchases made with Amazon Rewards",
         y = "Total Spent, in USD"
         )
```

```{r another facet chart, by purchaser}
ggplot(df, aes(x = OrderDate, y = ItemTotal)) +
    geom_point(color = "darkorchid4") +
    facet_wrap(~ BuyerName)
    labs(title = "Amazon Spending by One Family",
         subtitle = "Includes purchases made with Amazon rewards",
         y = "Amount Spent",
         x = "Order Date (June 2006 - July 2021)") + theme_bw(base_size = 15)
```

```{r}
ggplot(df, aes(x = ItemTotal, y = SavingsPercent)) +
    geom_point(color = "darkorchid4") +
    facet_wrap(~ BuyerName)
    labs(title = "Amazon Spending by One Family",
         subtitle = "Includes purchases made with Amazon rewards",
         y = "Percent Off List Price",
         x = "Purchase Price in USD") + theme_bw(base_size = 15)
```

# from https://moderndive.com/5-regression.html
```{r add best fit line}
ggplot(df, aes(x = ItemTotal, y = SavingsAmount)) +
  geom_point() +
  labs(x = "Cost of Item", y = "Dollars Saved",
       title = "Relationship between product cost and money saved") +
  geom_smooth(method = "lm", se = FALSE)
```

# from https://a-little-book-of-r-for-time-series.readthedocs.io/en/latest/src/timeseries.html
```{r store the dataframe as a time series}
timeseries <- ts(df)
timeseries
```

# That was not what I wanted.  Try putting OrderDate into its own .csv file, read it into R store it as a time series
```{r import OrderDate}
orderdates <- read.csv('OrderDate_only.csv',stringsAsFactors = FALSE)
head(orderdates)
```

# from https://www.statmethods.net/input/dates.html
```{r convert to dates}
dates <- as.Date(orderdates$OrderDate, "%m/%d/%Y")
head(dates)
```

```{r store dates as a time series}
ordertimeseries <- ts(dates)
head(ordertimeseries)
```

```{r try adding frequency}
ts_with_frequency <- ts(dates, start=c(2006,7), frequency = 365)
ts_with_frequency
# Ouch! This assumes that each order is a single month and stretches the time series out into the future!
```

# from https://www.tutorialspoint.com/how-to-convert-a-data-frame-column-to-date-that-contains-integer-values-in-r
```{r import OrderDate and change OrderDate column to date type}
orderdates_try <- read.csv('OrderDate_only.csv',stringsAsFactors = FALSE)
dates_as_dates <- data.frame(orderdates_try)
dates_as_dates <- transform(dates_as_dates, OrderDate=as.Date(as.character(OrderDate), "%Y%m%d"))
dates_as_dates             # Why does it change the date to 'NA?'
```

# from https://moderndive.com/5-regression.html
```{r correlation coefficient}
evals_ch6 <- df %>%
  select(ID, SavingsPercent, Category, BuyerName, ItemTotal)
glimpse(evals_ch6)
```

```{r}
evals_ch6 %>%
  get_correlation(formula = SavingsPercent ~ ItemTotal)
```

# from https://moderndive.com/6-multiple-regression.html
```{r add BuyerName}
ggplot(evals_ch6, aes(x = ItemTotal, y = SavingsPercent, color = BuyerName)) +
  geom_point() +
  labs(x = "Item Total in USD", y = "Percent Savings", color = "Purchaser") +
  geom_smooth(method = "lm", se = FALSE)
```

```{r histogram for outliers}
ggplot(df, aes(x = ItemTotal)) +
  geom_histogram(binwidth = 5, boundary = 0.4, color = "white") +
  labs(x = "Item Cost", 
       title = "Distribution of Items Purchased, n=2020") 
```

# from http://rstudio-pubs-static.s3.amazonaws.com/374830_ab4ac8951c94411aa9abfe5f561cef91.html
```{r time series data}
ggplot(df, aes(OrderDate, ItemTotal)) + geom_line() + ylab("Item Cost in USD") +
  xlab("date")
```

# from https://statisticsglobe.com/error-discrete-value-supplied-to-continuous-scale-in-r
```{r have to convert DayOfWeek to categorical first}
data_new <- df[df$Quantity != "8+", ]
ggplot(data_new, aes(y=Quantity, x = DayOfWeek)) + 
  geom_boxplot(aes(group=DayOfWeek)) + 
  #geom_point() +
  #scale_x_continuous(breaks=seq(0,7,1)) + 
  ggtitle("Order Count by Day")
```


```{r}
ggplot(df, aes(OrderDate, Quantity)) + geom_line() + ylab("Count of Items Ordered") +
  xlab("date")
```

# from https://otexts.com/fpp2/forecasting-decomposition.html
```{r}
fit <- stl(df, t.window=13, s.window="periodic",
  robust=TRUE)
fit %>% seasadj() %>% naive() %>%
  autoplot() + ylab("New orders index") +
  ggtitle("Naive forecasts of seasonally adjusted data")
```

# from https://youtu.be/UekOBfpu8m8
```{r try ARIMA}
data1 <- import("cl_Amazon_both_for_ARIMA.csv")

data1$DATES <- anydate(data1$OrderDate)

ggplot(data1, aes(DATES, Quantity)) + geom_line() + scale_x_date('Month') + ylab("Count of Purchases") + xlab("")

```

#This did not work
# from https://otexts.com/fpp2/arima-r.html
```{r}
data3 <- read.csv('unique_order_dates.csv',stringsAsFactors = FALSE)
data3_ts <- ts(data3)
```

```{r}
data3_ts %>% stl(s.window='periodic') %>% seasadj() -> eeadj
autoplot(eeadj)
```

## try using padr library
# from https://cran.r-project.org/web/packages/padr/vignettes/padr.html
# trying to fill in dates that have no entries; 'padding' the time series?
```{r}
install.packages("padr")
library(padr)
```

```{r}
 data1 %>%                     
  thicken('day') %>%
  group_by(OrderDate) %>%
  summarise(item_total = sum(ItemTotal)) %>%
  pad() %>%
  fill_by_value() %>%
  ggplot(aes(OrderDate, item_total)) + geom_line()
```

```{r convert column to dates in data3, then try and get weekly intervals with thicken()}
#UniqueOrderDate <- as.Date(data3$UniqueOrderDate, "%m/%d/%Y")   # doesn't work
data3$UniqueOrderDate <- as.Date(data3$UniqueOrderDate)
head(data3)
```

```{r use thicken() to try to get to weekly data intervals}
data4 <- data3 %>% thicken('week')
data4$OrderDate %>% get_interval()
# so this failure has something to do with a C++ wrapper from 15 years ago!  The 'fix' is atrocious, and I'm sure I'll break something. https://github.com/RcppCore/Rcpp/issues/1081
```

# try purrr, from https://algotech.netlify.app/blog/purrr-operly-fitting-multiple-time-series-model/
```{r}
install.packages("magrittr")
library(magrittr)
install.packages("tidymodels")
library(tidymodels)
```

# try to fill in the missing dates, from https://community.rstudio.com/t/how-to-fill-missing-dates-in-dataset/70517/6
```{r already have ubridate, need tibble and tidyr}
install.packages("tibble")
install.packages("tidyr")
library(tibble)
library(tidyr)
```

```{r}
install.packages("dplyr")
library(dplyr)
```

```{r import data}
data5 <- read.csv('indiv_order_dates_for_Python.csv',stringsAsFactors = FALSE)
head(data5)
```

```{r change to a tibble}
tibble5 <- tibble(data5$OrderDate, data5$Quantity, data5$Combining_Duplicates) %>% 
  mutate(OrderDate = mdy(OrderDate))
print(tibble5)
```
# I am DONE trying to get this time series wrangled into a regular time series in R!
# Time to try Python!